\documentclass{article}

\def\subj {CSCE}
\def\npart {221}
\def\nterm {Spring}
\def\nyear {2019}
\def\ncourse {Data Structures and Algorithms}



\input{header}

\setlength\parindent{0pt}

\begin{document}
\maketitle
{\small
  \noindent\textbf{Arrays, Linked Lists, and Recursion}\\
  Conduction of Heat in a 1-Dimensional rod, boundary conditions, equilibrium temperature distribution, heat condition in 2 or 3 dimensions.\hspace*{\fill} [1,3]

  \vspace{10pt}
  \noindent\textbf{Introduction to Analysis of Algorithms}\\
  Linearity, heat equation with zero temperatures at finite ends, orthogonality of functions, Laplace's equation; solutions and qualitative properties.\hspace*{\fill} [4]

  \vspace{10pt}
  \noindent\textbf{Stacks and Queues}\\
  Statement of Convergence Theorem, Fourier cosine and sine series, term-by-term differentiation of Fourier series, term-by-term integration of Fourier series, complex form of Fourier series.\hspace*{\fill} [5]

  \vspace{10pt}
  \noindent\textbf{Vectors, Lists, and Sequences}\\
  Vertically vibrating string, boundary conditions, vibrating string with fixed ends, vibrating membrane, reflection and refraction of electromagnetic and acoustic sound waves.\hspace*{\fill} [6]

  \vspace{10pt}
  \noindent\textbf{Trees and Search Trees}\\
  Characteristics for first order wave equations, method of characteristics for first order PDEs, one-dimensional wave equation, a vibrating string of fixed length, many quasilinear PDEs, semi-infinite strings and reflections. \hspace*{\fill} [7,10]

  \vspace{10pt}
  \noindent\textbf{Priority Queues, Heaps}\\
  Heat equation on an infinite domain, Fourier transform pair, inverse Fourier transform, convolution theorem.\hspace*{\fill} [8]

  \vspace{10pt}
  \noindent\textbf{Dictionaries, Hashing}\\
  Green's functions for boundary value problems for ODEs, method of eigenvalue expansion, nonhomogeneous boundary conditions.\hspace*{\fill} [9]

  \vspace{10pt}
  \noindent\textbf{Sorting, Sets and Selection}\\
  Green's functions for boundary value problems for ODEs, method of eigenvalue expansion, nonhomogeneous boundary conditions.\hspace*{\fill} [11]

  \vspace{10pt}
  \noindent\textbf{Text Compression}\\
  Green's functions for boundary value problems for ODEs, method of eigenvalue expansion, nonhomogeneous boundary conditions.\hspace*{\fill} [12.4]

  \vspace{10pt}
  \noindent\textbf{Graphs}\\
  Green's functions for boundary value problems for ODEs, method of eigenvalue expansion, nonhomogeneous boundary conditions.\hspace*{\fill} [13]
  }

\tableofcontents


	\section{The Heat Equation}

\end{document}
